A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2020-07-28 Analysis/nat_stories_lms.Rmd
add nested model comparison, anovas

labelled_pre_error <- data_pre_error %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("pre_error.rds")

@@

 a <- log_lik(brm_ngram_interact)

 b <- log_lik(brm_grnn_interact)

 c <- log_lik(brm_txl_interact)

 get_harmonic(a)

 get_harmonic(b)

 get_harmonic(c)
```

@@

>>>>

labelled_pre_error <- data_pre_error %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  select(rt, subject, Word_In_Story_Num, word, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("pre_error.rds")

@@

 # a <- log_lik(brm_ngram_interact)
 # 
 # b <- log_lik(brm_grnn_interact)
 # 
 # c <- log_lik(brm_txl_interact)
 # 
 # get_harmonic(a)
 # 
 # get_harmonic(b)
 # 
 # get_harmonic(c)
```

@@

# Frequentist LMs

```{r}
#mean by subject for each word
d_lm <- labelled_pre_error %>% group_by(word, txl_center, ngram_center, grnn_center, freq_center, length_center,
                                        past_c_txl, past_c_ngram, past_c_grnn, past_c_freq, past_c_length, Word_ID) %>% 
  summarize(mean_rt=mean(rt))
```

## Models

We want to be able to do nested model comparision, so we want models with 
- only length, frequency fx
- each 1 surprisal model as predictor
- all the surprisals

For now, no interactions between surprisal and others, and if we include a predictor, include both current and lagged of it.

no_surp <- lmer(mean_rt ~ freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

ngram_only <- lmer(mean_rt ~ ngram_center + past_c_ngram + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

grnn_only <- lmer(mean_rt ~ grnn_center + past_c_grnn + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

txl_only <- lmer(mean_rt ~ txl_center + past_c_txl + freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

all_surp <- lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   grnn_center + past_c_grnn +
                   txl_center + past_c_txl +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

ngram_grnn <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   grnn_center + past_c_grnn +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

ngram_txl <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   txl_center + past_c_txl +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

grnn_txl <-  lmer(mean_rt ~ grnn_center + past_c_grnn +
                   txl_center + past_c_txl +
                   freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

no_lag <-  lmer(mean_rt ~ ngram_center +
                   grnn_center +
                   txl_center + 
                   freq_center*length_center+(1|word), data=d_lm)

no_surp_lag <-  lmer(mean_rt ~ ngram_center +  
                   grnn_center + 
                   txl_center +                    freq_center*length_center+past_c_freq*past_c_length+(1|word), data=d_lm)

  only_surp_lag <-  lmer(mean_rt ~ ngram_center + past_c_ngram + 
                   grnn_center + past_c_grnn +
                   txl_center + past_c_txl +
                   freq_center*length_center+(1|word), data=d_lm)
```

anova(all_surp, txl_only)
anova(all_surp, grnn_only)
anova(all_surp, ngram_only)

anova(txl_only, no_surp)
anova(grnn_only, no_surp)
anova(ngram_only, no_surp)

anova(all_surp, ngram_grnn)
anova(all_surp, ngram_txl)
anova(all_surp, grnn_txl)

anova(ngram_grnn, ngram_only)
anova(ngram_grnn, grnn_only)
anova(ngram_txl, ngram_only)
anova(ngram_txl, txl_only)
anova(grnn_txl, grnn_only)
anova(grnn_txl, txl_only)

anova(all_surp, no_lag)

anova(all_surp, no_surp_lag)

anova(no_surp_lag, no_lag)

anova(all_surp, only_surp_lag)

anova(only_surp_lag, no_lag)
```
In general, adding to the model makes it better.

The model with all 3 surprisal predictors is better than any model with only one. Any one surprisal predictor is better than no surprisals predictors.

However, adding ngram predictors to a model that already has txl & grnn does not help. 
In other cases, adding the 3rd surprisal source to the other two does help.

Ngram+Grnn is not better than Grnn only. Otherwise, pairs are better than singletons. This suggests that Ngram's info is a subset of GRNN, but not a subset of TXL.

Past surprisal predictors don't help (with or without past freq,length effects in the models), but past freq,length do (with or without past surprisal predictors).

### 2020-07-28 Papers/amlap2020_abstract.tex
new gams
add nested model comparison, anovas

% The next bit is for reserving sufficient space for authors,
% affiliations, and e-mail address.  No need to change for initial
% anonymous version.  For the final version, replace the
% \toggletrue{anonymous} with \togglefalse{anonymous} to de-anonymize.
\usepackage{etoolbox}
\newtoggle{anonymous}
\toggletrue{anonymous}

@@

\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
\authors{Veronica Boyce (MIT), Roger Levy (MIT)} %TODO, check what affiliation I have at conference time
\email{vboyce@mit.edu}
\newline
%
% Here goes the main text of your abstract:
%Comment: I feel like you could make a bigger deal of the current-word previous-word / RT correlation. Given how there are spillover effects for SPR and also for eye-tracking (if I’m not mistaken), it’s crazy that the relationship between RT/Surprisal for the previous word is basically flat, when it comes to Maze!
%Comment: I’m confused with Table 1: Doesn’t length have a bigger effect size than surprisal? On a related note: does it matter which has a bigger effect? Why is it important to say that surprisal has a bigger effect, compared to just a robust, significant effect?

@@

We also investigated the relationship between a word's predictability and its RT. We fit Bayesian regression models using suprisal, frequency, length, and surprisal:length and frequency:length interactions at both the current and previous word as predictors of RT (see Table~1 for details). We use 3 models for surprisals: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al. (2018), and a transformer model (TXL) from Dai et al. (2019). Across all models, we see large effects of current word surprisal and length, which noticeably exceed effects of previous word predictors (Table~1). We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3). The linear relationship matches known effects from eye-tracking and self-paced reading (Smith \& Levy 2013), but the lack of spillover indicates a potential advantage of Maze over eye-tracking and self-paced reading.

In summary, we provide an adaptation of the Maze task for long naturalistic texts, and show that participants can comprehend what they read during A-maze. Additionally, we find robust, localized surprisal effects, supporting Forster et al.'s (2009) argument that Maze forces highly incremental processing. Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading.

Our code and data are available at [link to repo redacted for anonymity].

@@

\vspace{.1em}
	\begin{small}{
			\noindent\textbf{References:}
			Boyce et al (2020). \textit{J Mem. Lang.} $\bullet$
			Dai et al (2019) arXiv:1901.02860 $\bullet$
			Forster et al (2009). \textit{Behav. Res. Methods} $\bullet$
			Futrell et al (2017). arXiv:1708.05763 $\bullet$
			Gulordava et al (2018).  \textit{NAACL-HLT 2018} $\bullet$
			Sloggett et al (2020). \textit{CUNY 2020} $\bullet$
			Smith \& Levy (2013). \textit{Cognition} $\bullet$
			Witzel et al (2012).  \textit{J Psycholinguist Res} 
			%TODO add cites for length, freq, surprisal justification
		}
	\end{small}

>>>>

% The next bit is for reserving sufficient space for authors,
% affiliations, and e-mail address.  No need to change for initial
% anonymous version.  For the final version, replace the
% \toggletrue{anonymous} with \togglefalse{anonymous} to de-anonymize.
\usepackage{etoolbox}
\newtoggle{anonymous}
\togglefalse{anonymous}

@@

\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
\authors{Veronica Boyce (MIT/Stanford), Roger Levy (MIT)} %TODO, check what affiliation I have at conference time
\email{vboyce@stanford.edu}
\newline
%
% Here goes the main text of your abstract:
%Comment: I feel like you could make a bigger deal of the current-word previous-word / RT correlation. Given how there are spillover effects for SPR and also for eye-tracking (if I’m not mistaken), it’s crazy that the relationship between RT/Surprisal for the previous word is basically flat, when it comes to Maze!
%Comment: I’m confused with Table 1: Doesn’t length have a bigger effect size than surprisal? On a related note: does it matter which has a bigger effect? Why is it important to say that surprisal has a bigger effect, compared to just a robust, significant effect?

@@

We also investigated the relationship between a word's predictability and its RT. We fit Bayesian regression models using suprisal, frequency, length, and surprisal:length and frequency:length interactions at both the current and previous word as predictors of RT (see Table~1 for details). We use 3 models for surprisals: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al. (2018), and a transformer model (TXL) from Dai et al. (2019). Across all models, we see large effects of current word surprisal and length, which noticeably exceed effects of previous word predictors (Table~1). We see minimal to no frequency effects (consistent with Shain 2019). In a nested model comparison, we found GRNN and TXL encoded complementary information, but adding 5-gram surprisals did not improve a regression that already had GRNN. We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3). The linear relationship matches known effects from eye-tracking and self-paced reading (Smith \& Levy 2013), but the lack of spillover indicates a potential advantage of Maze over eye-tracking and self-paced reading.

In summary, we provide an adaptation of the Maze task for long naturalistic texts, and show that participants can comprehend what they read during A-maze. Additionally, we find robust, localized surprisal effects, supporting Forster et al.'s (2009) argument that Maze forces highly incremental processing. Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading. Code and data at github.com/vboyce/amaze-natural-stories.

@@

\vspace{.1em}
	\begin{small}{
			\noindent\textbf{References:}
			Boyce et al (2020). \textit{J Mem. Lang.} $\bullet$
			Dai et al (2019) arXiv:1901.02860 $\bullet$
			Forster et al (2009). \textit{Behav. Res. Methods} $\bullet$
			Futrell et al (2017). arXiv:1708.05763 $\bullet$
			Gulordava et al (2018).  \textit{NAACL-HLT 2018} $\bullet$
			Shain (2019). \textit{NAACL-HLT 2019} $\bullet$
			Sloggett et al (2020). \textit{CUNY 2020} $\bullet$
			Smith \& Levy (2013). \textit{Cognition} $\bullet$
			Witzel et al (2012).  \textit{J Psycholinguist Res.} 
		}
	\end{small}
